Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-03-03 paper/main.tex
small changes in section 3.

With this lightweight version of $\mcycle(\vec{e})$, the formula
$\Phi_{C,K}$ actually says that
for every nonempty set
$T\subseteq\delta$, $T$ satisfies $\varphi$ if and only if $\rem(T)$ satisfies
$\psi$. Note that the formula is not affected by the placement of
transitions in the automaton. The only aspect of a transition $t$
reflected by the formula is its set of marks $\mks(t)$. Hence, we do
not have to distinguish between transitions with the same sets of
marks.

@@

The lemma suggests the following simplification of the whole formula
$\Phi_{C,K}$ built with $\mcycle_1(\vec{e})$. Before we build the
formula, we compute the equivalence $\sim$ on automata transitions and
remove all transitions except one of each class. Then we build the
formula for this pruned automaton as described above. The more
transitions we removed, the shorter formula with less $e_t$ variables
we obtain. If the formula $\Phi_{C,K}$ for the pruned automaton is
satisfiable and we derive $\rem$ and $\psi$ from its model, we can use
safely them also on the original automaton. We only need to extend
$\rem$ to all transitions such that it changes the acceptance marks on
all equivalent transitions in the same way.

@@

The formula $\mcycle_3(\vec{e})$ says that $\Te$ is nonempty and each
set $S\subseteq Q$ satisfies (\ref{en:1}) or (\ref{en:2}) or
(\ref{en:3}).  For each state $q\in Q$, variable $s_q$ says whether
$q$ is in the current set $S$ or not.
\[s_q=\left\{
    \begin{array}{ll}
      1 & \textrm{if }q\in S\\
      0 & \textrm{otherwise}    
    \end{array}\right.
\]
By $\vec{s}$ we denote the vectors of all variables of the form $s_q$.
The formula $\mcycle_3(\vec{e})$ is defined as follows.
\begin{align*}
  \mcycle_3(\vec{e})
  &=\bigvee_{t\in\delta}e_t~\wedge~\forall\vec{s}~.~\zeta_{\ref{en:1}}(\vec{e},\vec{s})\,\vee\,
    \zeta_{\ref{en:2}}(\vec{e},\vec{s})\,\vee\,
    \zeta_{\ref{en:3}}(\vec{e},\vec{s})\\ %\qquad\textrm{where}\\
  \zeta_{\ref{en:1}}(\vec{e},\vec{s}) &= \bigwedge_{t=(p,a,M',q)\in\delta}\hspace{-5mm}\big(e_t\implies (s_p\wedge s_q)\big)\\
  \zeta_{\ref{en:2}}(\vec{e},\vec{s}) &= \bigwedge_{t=(p,a,M',q)\in\delta}\hspace{-5mm}\big(e_t\implies (\neg s_p\wedge \neg s_q)\big)\\
  \zeta_{\ref{en:3}}(\vec{e},\vec{s}) &= \big(\bigvee_{t=(p,a,M',q)\in\delta}\hspace{-5mm}e_t\wedge s_p\wedge \neg s_q\big) ~\wedge~
  \big(\bigvee_{t=(p,a,M',q)\in\delta}\hspace{-5mm}e_t\wedge \neg s_p\wedge s_q\big)
\end{align*}
\todo{Zminit, ze z univerzalniho kvantifikatoru se stane existencni.}

>>>>

The formula $\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$ actually
says that for every nonempty set $T\subseteq\delta$, $T$ satisfies
$\varphi$ if and only if $\rem(T)$ satisfies $\psi$. Note that the
formula is not affected by the placement of transitions in the
automaton. The only aspect of a transition $t$ reflected by the
formula is its set of marks $\mks(t)$. Hence, we do not have to
distinguish between transitions with the same sets of marks.

@@

The lemma suggests the following simplification of the whole formula
$\Phi_{1,C,K}$ built with $\mcycle_1(\vec{e})$. Before we build the
formula, we compute the equivalence $\sim$ on automata transitions and
remove all transitions except one of each class. Then we build the
formula for this pruned automaton as described above. The more
transitions we removed, the shorter formula with less $e_t$ variables
we obtain. If the formula $\Phi_{1,C,K}$ for the pruned automaton is
satisfiable and we derive $\rem$ and $\psi$ from its model, we can use
safely them also on the original automaton. We only need to extend
$\rem$ to all transitions such that it changes the acceptance marks on
all equivalent transitions in the same way.

@@

The formula $\mcycle_3(\vec{e})$ says that $\Te$ is nonempty and each
set $S\subseteq Q$ satisfies (\ref{en:1}) or (\ref{en:2}) or
(\ref{en:3}).  For each state $q\in Q$, variable $s_q$ says whether
$q$ is in the current set $S$ or not.
\[s_q=\left\{
    \begin{array}{ll}
      1 & \textrm{if }q\in S\\
      0 & \textrm{otherwise}    
    \end{array}\right.
\]
By $\vec{s}$ we denote the vectors of all variables of the form $s_q$.
The formula $\mcycle_3(\vec{e})$ is defined as follows.
\begin{align*}
  \mcycle_3(\vec{e})
  &=\bigvee_{t\in\delta}e_t~\wedge~\forall\vec{s}~.~\zeta_{\ref{en:1}}(\vec{e},\vec{s})\,\vee\,
    \zeta_{\ref{en:2}}(\vec{e},\vec{s})\,\vee\,
    \zeta_{\ref{en:3}}(\vec{e},\vec{s})\\ %\qquad\textrm{where}\\
  \zeta_{\ref{en:1}}(\vec{e},\vec{s}) &= \bigwedge_{t=(p,a,M',q)\in\delta}\hspace{-5mm}\big(e_t\implies (s_p\wedge s_q)\big)\\
  \zeta_{\ref{en:2}}(\vec{e},\vec{s}) &= \bigwedge_{t=(p,a,M',q)\in\delta}\hspace{-5mm}\big(e_t\implies (\neg s_p\wedge \neg s_q)\big)\\
  \zeta_{\ref{en:3}}(\vec{e},\vec{s}) &= \big(\bigvee_{t=(p,a,M',q)\in\delta}\hspace{-5mm}e_t\wedge s_p\wedge \neg s_q\big) ~\wedge~
  \big(\bigvee_{t=(p,a,M',q)\in\delta}\hspace{-5mm}e_t\wedge \neg s_p\wedge s_q\big)
\end{align*}
